---
title: "fieldtrialcheck"
author: "Opamelas83"
date: "2022-08-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction

Sources of variability
Fixed Factors:
blockNumber: fixed-effect for block replicate effects with levels uniquely nested within trials combination
Plantharvested: cov  

Random Factors: germplasmName: random genetic effect, primary aim is to compute BLUPs for this term

Error Variance:
studyName: heterogeneous error variances for each trial

```{r}
library(workflowr); library(tidyverse); library(here); library(lubridate);  library(lme4); library(magrittr)
colClasses <- readr::read_csv(here::here("data/IITA2021_CABASE", "colClasses.csv"))
colString <- paste0(substring(colClasses$colClass, 1, 1), collapse="")
Trialcheck1 <- readr::read_csv(here::here("data/IITA2021_CABASE", "UYT_GXE_pheno.csv"), col_types=colString)
Trialcheck1 <- Trialcheck1 %>% filter(!(studyYear == 2020))
```

#create explicitly nested design variables (yearInLoc, trialInLocYr, repInTrial, blockInRep).
```{r}
#Trialcheck1 %<>% 
 #    group_by(studyName) %>% 
   #  summarize(N_replicate=length(unique(replicate)),
        #       N_blockNumber=length(unique(blockNumber)),
         #      doRepsEqualBlocks=all(replicate==blockNumber)) %>% 
    # ungroup() %>% 
     #mutate(CompleteBlocks=ifelse(N_replicate>1,TRUE,FALSE),
      #      IncompleteBlocks=ifelse(N_blockNumber>1 & !doRepsEqualBlocks,TRUE,FALSE)) %>% 
     #left_join(Trialcheck1) %>% 
    # mutate(yearInLoc=paste0(programName,"_",locationName,"_",studyYear),
          #  trialInLocYr=paste0(yearInLoc,"_",studyName),
           # repInTrial=paste0(trialInLocYr,"_",replicate))
```

Cassavabase downloads use very long column-names corresponding to the full trait-ontology name. For convenience, I replace these names with abbreviations, documented here.
I also use this opportunity to subselect traits.
```{r}
Trialcheck2 <- Trialcheck1 %>%
  rename(CMD1S ="cassava mosaic disease severity 1-month evaluation|CO_334:0000191",
        CMD3S = "cassava mosaic disease severity 3-month evaluation|CO_334:0000192",
        DMC = "dry matter content percentage|CO_334:0000092",
        FRWT = "fresh storage root weight per plot|CO_334:0000012",
        PHar = "plant stands harvested counting|CO_334:0000010",
        FTWT = "fresh shoot weight measurement in kg per plot|CO_334:0000016") %>% 
  mutate(yearInLoc = paste0(locationName,"_",studyYear),
         PlantsPerPlot = max(PHar, na.rm=TRUE),
         PropPHar = PHar/PlantsPerPlot,
         repInTrial = paste0(studyYear,"_",replicate),
         blockInRep = paste0(repInTrial,"_",blockNumber)) %>%
  dplyr::select(studyName,
         yearInLoc ,
         repInTrial,
         blockInRep,
         germplasmName,
         CMD1S,
         CMD3S,
         DMC,
         FRWT,
         PropPHar, 
         PHar,
         PlantsPerPlot,
         FTWT)
```


Variance component estimates
```{r}
TRaits <- c("FRWT", "FTWT")#, "CMD1S", "CMD3S") 
PHenos <-Trialcheck2 %>% 
      # Convert the data to "long format" 
      pivot_longer(cols = all_of(TRaits), 
                   names_to = "TRait", values_to = "Value") %>%
      # Remove missing values
      filter(!is.na(Value))

PHenos <- PHenos %>% mutate(studyName = as.factor(studyName),
                            yearInLoc = as.factor(yearInLoc),
                            repInTrial = as.factor(repInTrial),
                            blockInRep = as.factor(blockInRep),
                            germplasmName = as.factor(germplasmName),
                            PropPHar = as.factor(PropPHar)) 
WW <- PHenos %>% group_by(studyName) %>% summarise(n=length(unique(PropPHar))) %>% filter(n==1)
WWs <- droplevels(WW$studyName) %>% table %>% names
TRials <- PHenos$studyName %>% table %>% .[order(.)] %>% names

j <- TRaits[1]
i <- TRials[1]
results <- tibble()

for(j in TRaits){
  for(i in TRials){
    DataM <- PHenos %>% filter(studyName == i, TRait == j)
    meanNreps <- DataM %>%
     count(germplasmName) %$% mean(n)
    Formula1 <- "Value ~ (1|repInTrial) + (1|germplasmName)" 
Formula <- ifelse(unique(DataM$studyName) %in% WWs, Formula1, paste0(Formula1," + PropPHar"))
    if(nrow(DataM) != 0){
          try(MixedModels <-lmer(as.formula(Formula), data = DataM))
     # try(MixedModelsrd <- lm(Value ~ PropPHar + repInTrial, data = DataM))
    try(result <- tibble(Trait = j,
                         Trial = i,
                         VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
                         VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[4,2],
                         H2 = VarG/(VarG + VarE/meanNreps),
                         Rel =  suppressWarnings(MuMIn::r.squaredGLMM(MixedModels)[2]),
                         model = Formula))
                         #DvPrb = anova(MixedModels, MixedModelsrd) %>% as.data.frame() %>% .[2,"Pr(>Chisq)"] %>% as.numeric))
    try(results <- bind_rows(results, result))
    #rm(MixedModels); rm(result); rm(MixedModelsrd)
    }
    }
}
results


write.table(results, file = "Result/MixedModelResults.csv", quote = F, sep =",",
            row.names = F)

#results %>% group_by(Trial) %>%
  #summarise(Ntraits = n(),
  #          VarGmean = mean(VarG, na.rm = F),
   #         VarEmean = mean(VarE, na.rm = F),
    #        H2mean = mean(H2, na.rm = F),
     #       Real = mean(Real, na.rm = F),
      #      DvPrb = mean(DvPrb, na.rm = F)) -> results2

#write.table(results2, file = "Result/MeansVarCompMixedModels.csv", quote = F,
        #    sep = ",", row.names = F)
```


```{r}
library(reactable)
```

Work with only two traits (without disease)
```{r}
results <- read_csv("Result/MixedModelResults.csv")
results$Location <- results$Trial
results$Location[grep(pattern = "AB", x = results$Trial)] <- "ABUJA"
results$Location[grep(pattern = "AG", x = results$Trial)] <- "AGO"
results$Location[grep(pattern = "IB", x = results$Trial)] <- "IBADAN"
results$Location[grep(pattern = "IK", x = results$Trial)] <- "IKENNE"
results$Location[grep(pattern = "KN", x = results$Trial)] <- "KANO"
results$Location[grep(pattern = "MK", x = results$Trial)] <- "MOKWA"
results$Location[grep(pattern = "ON", x = results$Trial)] <- "ONNE"
results$Location[grep(pattern = "OT", x = results$Trial)] <- "OTOBI"
results$Location[grep(pattern = "UB", x = results$Trial)] <- "UBIAJA"
results$Location[grep(pattern = "UM", x = results$Trial)] <- "UMUDIKE"
results$Location[grep(pattern = "ZA", x = results$Trial)] <- "ZARIA"

resultsW <- results %>% mutate(Mtrial = if_else((str_detect(results$Trial, "^*19") | (str_detect(results$Trial, "^*18") & Location == "AGO")), paste0(str_extract(Trial, "\\d{2}+"),"_",substr(Trial, 17, 20),"_",Location), paste0(str_extract(Trial, "\\d{2}+"),"_",substr(Trial, 10, 13),"_",Location)))

 

ggplot(resultsW) + 
  geom_text(aes(x = H2, y = Rel, label = Mtrial), size = 2) + 
  xlab("Heritability") +
  ylab("Reliability")
```

#Filter my database to eliminate the trials with H2 < 0.25 and Reliability < 40% 
```{r}
resultstrial <- results %>% 
  filter(H2 > 0.25 & Rel > 0.4) %>%
  dplyr::select(Trial)
undesiredtrials <- results%>%
  filter(H2 < 0.25 | Rel < 0.4)  %>%
  dplyr::select(Trial)
#I should remove also trials common to both groups that means:  if a  trial has a bad genetic value for any of the trait, it will not be considered. If set A trial has bee removed, the set B should be too because I would keep the same rep for the 67 clones. 
undesired <- undesiredtrials$Trial
Trialcheckfiltered <- Trialcheck1 %>% filter(!studyName %in% undesired) #This should be my final trial I used to run the model.
write.table(Trialcheckfiltered, file = "data/IITA2021_CABASE/UYT_GXE_phenofiltered.csv")#, quote = F, sep = ",", row.names = F)
undesired_2022 <- c("17.GS.C2.setB.UYT.36.AG",
  "18.GS.C2.setA.UYT.36.KN",
  "18.GS.C2.setB.UYT.36.KN",
  "18.GS.C2.setB.UYT.36.OT",
  "18.GS.C2.UYT.36.setA.AG",
  "19.GS.C2.UYT.36.setA.MK",
  "19.GS.C2.UYT.36.setA.ZA",
  "19.GS.C2.UYT.36.setB.AG",
  "19.GS.C2.UYT.36.setB.ON",
 "19.GS.C2.UYT.36.setB.ZA",
 "20.GS.C2.UYT.36.setB.IB") 
# I extracted my 28 trials I used to run the model. 
Trialcheckfiltered_2022 <- resultsW %>% 
  filter(!Trial %in% undesired_2022) %>%  
  filter(H2 > 0.25 & Rel > 0.4)                                      

ggplot(Trialcheckfiltered_2022) + 
  geom_text(aes(x = H2, y = Rel, label = Mtrial, color = Trait), size = 2) +
  xlab("Heritability") +
  ylab("Reliability")
```


Luc
```{r}

#Trialcheck1 <- readr::read_csv(here::here("data/IITA2021_CABASE", "UYT_GXE_pheno.csv"), col_types=colString)

#Trials <- Trialcheck1$studyName %>% table %>% .[order(.)] 

#Trialcheck2 <- Trialcheck1 %>% filter(studyName == Trials[1]) %>% 
 # rename(FRY = "fresh storage root weight per plot|CO_334:0000012",
 #        PHar = "plant stands harvested counting|CO_334:0000010",
  #       FTY = "fresh shoot weight measurement in kg per plot|CO_334:0000016") %>% 
 # mutate(blockNumber = as.factor(blockNumber)) %>% 
  #lmer(formula = FRY ~ blockNumber + (1|germplasmName), data = .)
#visualize ()
#anova(Trialcheck2)

#Trialcheck2rd <- Trialcheck1 %>% filter(studyName == Trials[1]) %>% 
 # rename(FRY = "fresh storage root weight per plot|CO_334:0000012",
 #        PHar = "plant stands harvested counting|CO_334:0000010") %>% 
  # mutate(blockNumber = as.factor(blockNumber)) %>% 
  #  lmer(formula =  FRY ~ PHar, data = .)#blockNumber
#anova(Trialcheck2, Trialcheck2rd)
#results <- tibble()

#for(i in Trials){
#  try(MixedModels <- Trialcheck1 %>% filter(studyName == i) %>% 
 # rename(FRY = "fresh storage root weight per plot|CO_334:0000012",
 #        PHar = "plant stands harvested counting|CO_334:0000010") %>% 
  #mutate(blockNumber = as.factor(blockNumber)) %>% 
  #  lmer(formula =  FRY ~ PHar + (1|germplasmName), data = .))#blockNumber#
#  try(MixedModelsrd <- Trialcheck1 %>% filter(studyName == i) %>% 
 # rename(FRY = "fresh storage root weight per plot|CO_334:0000012",
 #        PHar = "plant stands harvested counting|CO_334:0000010") %>% 
 # mutate(blockNumber = as.factor(blockNumber)) %>% 
  #  lm(formula =  FRY ~ PHar, data = .))#blockNumber
  
#  try(result <- tibble(Trial = i,
 #                      VarG = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% .[1,2],
 #                      VarE = as.data.frame(VarCorr(MixedModels))[,c("grp","vcov")] %>% #.[2,2],
  #                     H2 = VarG/(VarG + VarE),
  #                     Real = suppressWarnings(MuMIn::r.squaredGLMM(MixedModels)[2]),
   #   DvPrb = anova(MixedModels, MixedModelsrd) %>% as.data.frame() %>% 
   #     .[2,"Pr(>Chisq)"] %>% as.numeric))
 # try(results <- rbind(results, result))
# rm(MixedModels); rm(result); rm(MixedModelsrd)
#}
#results

```


