---
title: "MakeCSCfile"
author: ""
date: "2021-10-17"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Introduction
 1- I would like to write 67 new .CSC file with "TMEB419.CSC like template and named with the cultivar name Culf$"CNAME"
 2- Then in each new file, replace "CSUC0024 TMEB419" by the cultivar ID = Culf$"INGENO""+ Culf$"CNAME"
 3- Also write the colum "TRTNO" the corresponding numbe "@C" from Culf 
 
 
 ### Load packages 
```{r load packages, message=FALSE}
ip <- installed.packages()
packages_used <- c("workflowr", "tidyverse", "DSSAT", "here", "lubridate")
for (package in packages_used){
  if (!(package %in% ip[,"Package"])) install.packages(package)
}
library(tidyverse)library(tidyverse)
library(zoo)
library(rsq)
```

 
Make my Culf a part of the original "UYAB1901.CSX" file.
```{r}
Culf <- read.table("UYUB1901.CSX", skip = 81, header = T, nrows = 67, sep = "")
#Varieties <- Varieties$CNAME
#head(Varieties)
#CulfTest <- read_lines(Varieties)
#Culf <- CulfTest
```

Obtain the Trial names
```{r}
library(data.table)
TrialList <- dir()[dir()%like%".CSX"]
```

Make file name with the colum "CNAME" Cul. file
```{r}
makeFileName <- function(X){
  Name <- X$"CNAME"
  #CulName <- paste0(Name)
  #return(list(Name, CulName))
  return(Name)
} 
```

Make the cultivar ID with the first+the second colum from the CUL. file 
```{r}
makeCulID <- function(X){
  ID1 <- X$"INGENO"
  Name <- X$"CNAME"
  CulID <- paste0(ID1," ", Name)
  return(CulID)
} 
```

Make traitement rows
```{r}
makeTrtRow <- function(culNum, x){
    rowEnd <- "     0      0      0      0"
    cn <- paste0(ifelse(culNum > 9, "", " "), culNum)
    pad <- paste0(rep(" ", 85), collapse="")
    return(paste0(x, pad, cn, rowEnd))
  }
  
```

Write file.CSC, 
```{r}
#rowEnd <- "  1  0  0  1  0  0  0  0  0  0  1  1"
#varNameSpace <- 26
for(i in Culf$CNAME) {
    Variety <- i
    template <- "TMEB419.CSC"
    CSCFile <- read_lines(here::here("data/DSSATtempfile", template))
    fileName <- makeFileName(Culf[Culf$CNAME%in%Variety, ])
    FileID <- makeCulID (Culf[Culf$CNAME%in%Variety, ])
    CSCFile <- gsub("CSUC0024 TMEB419", paste0("CS", FileID), CSCFile, fixed=T)#write the new cultivar ID in my CSC File
    CSCTrialInfo <- CSCFile[1:2]
    for(x in TrialList){
      Trial <- read_lines(x)
      strtTrt <- grep(pattern = "*TREATMENTS", x = Trial, fixed=T)+2
      endTrt <- grep(pattern = "*CULTIVARS", x = Trial, fixed=T)-2
      Trial <- Trial[strtTrt:endTrt]
      Trial <- substr(Trial, start = 10, stop = 37)
      write.table(Trial, "test.txt", quote = F, col.names = F, row.names = F)
      Trial <- read.table("test.txt", header =F)
      colnames(Trial) <- c("Clone", "Treatment")
      Trial <- Trial[Trial$Clone%in%Variety,]
      head(Trial)
      CSCTrialInfo <- c(CSCTrialInfo, makeTrtRow(Trial$Treatment, x))
    
    }
    
    myfile <- here::here("data", "CSfile", paste0(fileName, ".CSC"))
    write_lines(x = CSCTrialInfo, file = myfile, append=F)
}
```
