---
title: "StatGXE"
author: "Opamelas83"
date: "2022-04-27"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## This analysis is to analyse the difference beteewn observed and simulated data from DSSAT OUTPUT. I try to see how the difference variation is going with statistical method
#Data preparation for Statistical analysis

```{r}
library(tidyverse)
library(statgenGxE)
library(corrplot)
library(gridExtra)
library(grid)
library(gridtext)
library(lemon)
GLUEVAL2 <- read_csv("data/GLUEVALana2.csv")
EVAL2StatGXE1 <- GLUEVAL2 %>% select(source, Clone, HWAMMS, HWAMMSsq, ELEV, LAT, LONG, RAIN, RAIN_6, RAIN_9, RAIN_12, RH2M, RH2M_6, RH2M_9, RH2M_12, WIND, WIND_6, WIND_9, WIND_12, SRAD, SRAD_6, SRAD_9, SRAD_12, TMIN, TMIN_6, TMIN_9, TMIN_12, TMAX, TMAX_6, TMAX_9, TMAX_12, PAR, PAR_6, PAR_9, PAR_12, year)
#EVAL2StatGXE1$HWAMMS <- multMissing(EVAL2StatGXE1$HWAMMS, maxIter = 10, naStrings = NULL)
#EVAL2StatGXE1$HWAMMSsq <- multMissing(EVAL2StatGXE1$HWAMMSsq, maxIter = 10, naStrings = NULL)

EVAL2StatGXE2 <- EVAL2StatGXE1 %>%
  filter(source %in% names(table(EVAL2StatGXE1$source))[table(EVAL2StatGXE1$source) > 20])## AMMI functions do not accept more than 30 % missing data. 

EVAL2StatGXE <- statgenSTA::createTD(data = EVAL2StatGXE2, genotype = "Clone", trial = "source")

```


# AMMI(Additive Main Effects and Multiplicative Interaction) analysis fits a model which involves the Additive Main effects (i.e. genotype and environment) along with Multiplicative Interaction
effects.
```{r}
AMMIStatGXE <- gxeAmmi(TD = EVAL2StatGXE, trait = "HWAMMS", nPC = 3)#byYear = 2019
summary(AMMIStatGXE)
write.csv(AMMIStatGXE$anova, "Result/AMMIStatGXEanova.csv", quote = F, row.names = F)
write.csv(AMMIStatGXE$importance, "Result/AMMIStatGXEimportance.csv", quote = F, row.names = F)

## Create an AMMI1 plot.
#genotype scaling, with emphasis on approximation of relations between genotypes (scale = 0)
#environment scaling, with emphasis on approximation of relations between environments (scale = 1)
#symmetric scaling, a compromise (scale = 0.5)
#jpeg(filename = "AMMI2env.jpeg", res = 400, width = 14, height = 8, units = "cm")
CFIG1 <- plot(AMMIStatGXE, plotType = "AMMI1", scale = 1, title = "Interactive PC1 vs main effects") + theme_classic() 
#dev.off()
CFIG2 <- plot(AMMIStatGXE, plotType = "AMMI2", scale = 0.5, title = "PC1 vs PC2")  + theme_classic
#dev.off()
nt <- theme(legend.position='none')
#p = list(CFIG1+nt, CFIG2+nt) %>% map(~.x + labs(x=NULL, y=NULL))

#jpeg(filename = "Result/Rcomparison.jpeg")
grid_arrange_shared_legend(CFIG1+nt, CFIG2+nt)#ncol = 2, nrow = 1
#dev.off() 
```

##I use the AMMI envScores to look for a link  with env parameters
```{r}
# weather data check
AMMIenvscores1 <- AMMIStatGXE$envScores # using for environ scores
AMMIenvscores2 <- as.data.frame(AMMIenvscores1) 
AMMIenvscores3 <- as_tibble(rownames_to_column(AMMIenvscores2)) %>% 
                          mutate(ENV = rowname) %>%
                                 #envScPC1 = PC1,
                                 #envScPC2 = PC2) %>%
                         select(-rowname)

AMMIenvMean1 <- AMMIStatGXE$envMean #using for environ mean (mean of root harvested by envi)n
AMMIenvMean2 <- as.data.frame(AMMIenvMean1) 
AMMIenvMean3 <- as_tibble(rownames_to_column(AMMIenvMean2)) %>% 
                          mutate(ENV = rowname,
                                envMean = AMMIenvMean1) %>% 
                         select(-rowname, -AMMIenvMean1)

EVAL2StatGXEenv <- EVAL2StatGXE1 %>% 
                     select(-Clone, -HWAMMS, -HWAMMSsq) %>%
                      mutate(ENV = source) %>%
                        dplyr::group_by(ENV) %>%
                         dplyr::summarise(ELEV = unique(ELEV), 
                                          LAT = unique(LAT), 
                                          LONG = unique(LONG), 
                                          RAIN = unique(RAIN),
                                          RAIN_6 = unique(RAIN_6), 
                                          RAIN_9 = unique(RAIN_9),
                                          RAIN_12 = unique(RAIN_12), 
                                          SRAD = unique(SRAD),
                                          SRAD_6 = unique(SRAD_6), 
                                          SRAD_9 = unique(SRAD_9), 
                                          SRAD_12 = unique(SRAD_12),  
                                          TMIN = unique(TMIN),
                                          TMIN_6 = unique(TMIN_6), 
                                          TMIN_9 = unique(TMIN_9), 
                                          TMIN_12 = unique(TMIN_12), 
                                          TMAX = unique(TMAX),
                                          TMAX_6 = unique(TMAX_6), 
                                          TMAX_9 = unique(TMAX_9), 
                                          TMAX_12 = unique(TMAX_12),
                                          RH2M = unique(RH2M),
                                          RH2M_6 = unique(RH2M_6), 
                                          RH2M_9 = unique(RH2M_9), 
                                          RH2M_12 = unique(RH2M_12), 
                                          WIND = unique(WIND),
                                          WIND_6 = unique(WIND_6), 
                                          WIND_9 = unique(WIND_9), 
                                          WIND_12 = unique(WIND_12),  
                                          PAR = unique(PAR),
                                          PAR_6 = unique(PAR_6), 
                                          PAR_9 = unique(PAR_9), 
                                          PAR_12 = unique(PAR_12))

StatGXEenv1 <- left_join(AMMIenvMean3, AMMIenvscores3)
StatGXEenv11 <- left_join(StatGXEenv1, EVAL2StatGXEenv)
StatGXEenv2 <- as_tibble(StatGXEenv11) %>% select(-ENV)
#corrplot(cor(StatGXEenv2[,-2]))
#corrplot(cor(StatGXEenv2[,-1]))
#jpeg(filename = "correlationEX.jpeg", res = 400, width = 14, height = 8, units = "cm")
StatGXEenv2_cor <- cor(StatGXEenv2)
StatGXEenv2_cor <- StatGXEenv2_cor[1:4,]
Figplot1 <- corrplot(StatGXEenv2_cor, tl.cex = 0.5, number.cex=0.5, method = "number", sig.level = 0.01, insig = "blank")
#dev.off()
```


#soil profile check
```{r}
Tral0 <- read_lines(here::here("SOIL.SOL"))
strtTrt01 <- grep(pattern = "*NG04041467", x = Tral0, fixed=T)

Tral01 <- read_lines(here::here("SOIL.SOL"), skip = 1953, n_max = 6 )#NG04365455 UMUDIKE
write_lines(Tral01, "test0.txt")
Tral01 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

Tral02 <- read_lines(here::here("SOIL.SOL"), skip = 1966, n_max = 6 )#NG04408648 ONNE
write_lines(Tral02, "test0.txt")
Tral02 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".") 

Tral03 <- read_lines(here::here("SOIL.SOL"), skip = 1979, n_max = 6 )#NG04309279 UBIAJA
write_lines(Tral03, "test0.txt")
Tral03 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")      

Tral04 <- read_lines(here::here("SOIL.SOL"), skip = 1992, n_max = 6 )#NG04304916 IKENNE
write_lines(Tral04, "test0.txt")
Tral04 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")      

Tral05 <- read_lines(here::here("SOIL.SOL"), skip = 2005, n_max = 6 )#NG04292017 OTOBI
write_lines(Tral05, "test0.txt")
Tral05 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")   

Tral06 <- read_lines(here::here("SOIL.SOL"), skip = 2018, n_max = 6 )#NG04283331 AGO-OWO
write_lines(Tral06, "test0.txt")
Tral06 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

Tral07 <- read_lines(here::here("SOIL.SOL"), skip = 2031, n_max = 6 )#NG04261714 IBADAN
write_lines(Tral07, "test0.txt")
Tral07 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

Tral08 <- read_lines(here::here("SOIL.SOL"), skip = 2044, n_max = 6 )#NG04171020 MOKWA 
write_lines(Tral08, "test0.txt")
Tral08 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

Tral09 <- read_lines(here::here("SOIL.SOL"), skip = 2057, n_max = 6 )#NG04184010 ABUJA 
write_lines(Tral09, "test0.txt")
Tral09 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

Tral10 <- read_lines(here::here("SOIL.SOL"), skip = 2070, n_max = 6 )#NG04088975 ZARIA 
write_lines(Tral10, "test0.txt")
Tral10 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

Tral11 <- read_lines(here::here("SOIL.SOL"), skip = 2083, n_max = 6 )#NG04041467 KANO 
write_lines(Tral11, "test0.txt")
Tral11 <- read.table("test0.txt", header = FALSE, sep = "", dec = ".")

TRAL <- bind_rows(Tral01, Tral02, Tral03, Tral04, Tral05, Tral06, Tral07, Tral08, Tral09, Tral10, Tral11, .id = "source")    
TRALA <- TRAL %>% mutate(Location = case_when(source == 1 ~ "UMUDIKE",
                                              source == 2 ~ "ONNE",
                                              source == 3 ~ "UBIAJA",
                                              source == 4 ~ "IKENNE",
                                              source == 5 ~ "OTOBI",
                                              source == 6 ~ "AGO-OWO",
                                              source == 7 ~ "IBADAN",
                                              source == 8 ~ "MOKWA",
                                              source == 9 ~ "ABUJA",
                                              source == 10 ~ "ZARIA",
                                              source == 11 ~ "KANO"),
                         Deth = V1,
                         SLLL_15 = case_when(V1 == 15 ~ V3),
                         SLLL_100 = case_when(V1 == 100 ~ V3),
                         SDUL_15 = case_when(V1 == 15 ~ V4),
                         SDUL_100 = case_when(V1 == 100 ~ V4),
                         SSAT_15 = case_when(V1 == 15 ~ V5),
                         SSAT_100 = case_when(V1 == 100 ~ V5),
                         SRGF_15 = case_when(V1 == 15 ~ V6),
                         SRGF_100 = case_when(V1 == 100 ~ V6),
                         SSKS_15 = case_when(V1 == 15 ~ V7),
                         SSKS_100 = case_when(V1 == 100 ~ V7),
                         SBDM_15 = case_when(V1 == 15 ~ V8),
                         SBDM_100 = case_when(V1 == 100 ~ V8),
                         SLOC_15 = case_when(V1 == 15 ~ V9),
                         SLOC_100 = case_when(V1 == 100 ~ V9),
                         SLCL_15 = case_when(V1 == 15 ~ V10),
                         SLCL_100 = case_when(V1 == 100 ~ V10),
                         SLSI_15 = case_when(V1 == 15 ~ V11),
                         SLSI_100 = case_when(V1 == 100 ~ V11),  
                         SLNI_15 = case_when(V1 == 15 ~ V13),
                         SLNI_100 = case_when(V1 == 100 ~ V13), 
                         SLHW_15 = case_when(V1 == 15 ~ V14),
                         SLHW_100 = case_when(V1 == 100 ~ V14),
                         SCEC_15 = case_when(V1 == 15 ~ V16),
                         SCEC_100 = case_when(V1 == 100 ~ V16)) %>%
  filter(V1 == 15 | V1 == 100) %>%
  select(-V1, -V2, -V3, -V4, -V5, -V6, -V7, -V8, -V9, -V10, -V11, -V12, -V13, -V14, -V15, -V16, -V17, -source)
  
 StatGXEenv1 <- left_join(AMMIenvMean3, AMMIenvscores3)
 StatGXEenvsol1 <- StatGXEenv1 %>% 
   mutate(Location = case_when(ENV == "Abuja19" ~ "ABUJA",
                               ENV == "Ago17" ~ "AGO-OWO",
                               ENV == "Ago18" ~  "AGO-OWO",
                               ENV == "Ago19" ~ "AGO-OWO",
                               ENV == "Ibadan17" ~ "IBADAN",
                               ENV == "Ibadan18" ~ "IBADAN", 
                               ENV == "Ibadan19" ~ "IBADAN",
                               ENV == "Ikenne17" ~ "IKENNE",
                               ENV == "Ikenne18" ~ "IKENNE",
                               ENV == "Ikenne19" ~ "IKENNE",
                               ENV == "Kano18" ~ "KANO",
                               ENV == "Mokwa17" ~ "MOKWA",
                               ENV == "Mokwa18" ~ "MOKWA",
                               ENV == "Mokwa19" ~ "MOKWA",
                               ENV == "Onne18" ~ "ONNE",
                               ENV == "Otobi18" ~ "OTOBI",
                               ENV == "Ubiaja19" ~ "UBIAJA", 
                              ENV == "Umudike18" ~ "UMUDIKE"))
                               
StatGXEenvsol2 <- left_join(StatGXEenvsol1 , TRALA) %>%
  select(-ENV, Deth)#-Location
StatGXEenvsol3 <- StatGXEenvsol2 %>% filter(!is.na(SLNI_15)) %>% 
  select(envMean, Location, PC1, PC2, PC3, SLLL_15, SDUL_15, SSAT_15, SSKS_15, SBDM_15, SLOC_15, SLCL_15, SLSI_15, SLHW_15, SCEC_15)#SRGF_15, SLNI_15,

StatGXEenvsol4 <- StatGXEenvsol2 %>% filter(!is.na(SLNI_100)) %>% 
  select(envMean, Location, PC1, PC2, PC3, SLLL_100, SSAT_100, SDUL_100, SSKS_100, SBDM_100, SLOC_100, SLCL_100, SLSI_100, SLHW_100, SCEC_100) #SLNI_100, SRGF_100,

StatGXEenvsol5 <- left_join(StatGXEenvsol3 , StatGXEenvsol4) %>% select(-Location) 
#jpeg(filename = "correlationENsol.jpeg", res = 400, width = 14, height = 8, units = "cm")
#Run only for 15cm of soil profile
StatGXEenvsol3 <- StatGXEenvsol3 %>% select(-Location)
StatGXEenvsol3_cor <- cor(StatGXEenvsol3)
StatGXEenvsol3_cor <- StatGXEenvsol3_cor[1:4, ]
Figplot2 <- corrplot(StatGXEenvsol3_cor, tl.cex = 0.5, number.cex=0.6, method = "number", sig.level = 0.01, insig = "blank")
#jpeg(filename = "Result/Rcomparison.jpeg")
#dev.off()
```

#plot all environments parameter in the same plot
```{r}
StatGXEenvT <- left_join(StatGXEenv2,StatGXEenvsol3)
StatGXEenvT_cor <- cor(StatGXEenvT)
StatGXEenvT_cor <- StatGXEenvT_cor[1:4,]
corrplot(StatGXEenvT_cor, tl.cex = 0.5, method = "number", sig.level = 0.01, insig = "blank")
```

No longer important   ##filter environ with total dry root mean between 1st and 3rd quantile to continue Model calibration
```{r}
#envMeanFiltered <- AMMIenvMean3 %>% filter(envMean > quantile(envMean, 0.25) & envMean < quantile(envMean, 0.75))
#genoMeanFiltered <- AMMIgenoMean3 %>% filter(genoMean > quantile(genoMean, 0.25) & genoMean < quantile(genoMean, 0.75))
```


# I use the AMMI genoScores to look for a link  with penotype  parameters

```{r}
PhenoDIS <- read_csv("data/IITA2021_CABASE/UYT_GXE_pheno.csv")
spec(PhenoDIS)
PhenoDISE <- PhenoDIS %>% rename(Year = studyYear,
               Clone = germplasmName,
               Location = locationName,
               CB_3Min = "cassava bacterial blight incidence 3-month evaluation|CO_334:0000178",
               CB_3Msev = "cassava bacterial blight severity 3-month evaluation|CO_334:0000175",
               CGM_1sev = "cassava green mite severity first evaluation|CO_334:0000189",
               CGM_2sev = "cassava green mite severity second evaluation|CO_334:0000190",
               CMO_1Min = "cassava mosaic disease incidence 1-month evaluation|CO_334:0000195",
               CMO_3Min = "cassava mosaic disease incidence 3-month evaluation|CO_334:0000196",
               CMO_1Msev = "cassava mosaic disease severity 1-month evaluation|CO_334:0000191",
               CMO_3Msev = "cassava mosaic disease severity 3-month evaluation|CO_334:0000192",
               HWAM = "fresh storage root weight per plot|CO_334:0000012",
               CWAM = "fresh shoot weight measurement in kg per plot|CO_334:0000016") %>%
            select(Clone, CB_3Min, CB_3Msev, CGM_1sev, CGM_2sev, CMO_1Min, CMO_3Min,                         CMO_1Msev, CMO_3Msev)# year, Location, HWAM, CWAM
write.csv(PhenoDISE, "data/IITA2021_Mich/PhenoDisea.csv", quote = F, row.names = F)
```

```{r}
#phenotype data measured from Ibadan 2021.
Phenocomp1 <- read_csv("data/IITA2021_Mich/Phenotyp_Mich2021.csv")
Phenocomp2 <- Phenocomp1  %>% 
        group_by(accession_name) %>%
        dplyr::summarise_all(funs(mean(., na.rm = TRUE)))
Phenocomp3 <- Phenocomp2 %>% mutate(Clone = accession_name,
                                    branchP = branching_point) %>%
            select(Clone, stem_nu, stem_D, branchP, plant_H)

#complete with leaf area data
Phenoleaf1 <- read_csv("data/IITA2021_Mich/AreaData.csv")
spec(Phenoleaf1)
#To solve Clone name issue
#Phenoleaf1$accession_name <- Phenoleaf1$accession_name %>%
 # gsub(pattern = "IITA_TMS_IB", replacement = "IITA-TMS-") %>% 
 # gsub(pattern = "TMS14F1159P00", replacement = "TMS14F1159P000")
Phenoleaf1$accession_name <- Phenoleaf1$accession_name %>% str_replace_all( c("IITA_TMS_IBA00070"="IITA-TMS-IBA000070",
                                             "IITA_TMS_IBA30572"="IITA-TMS-IBA30572",
                                             "IITA_TMS_IBA980581"="IITA-TMS-IBA980581",
                                             "IITA_TMS_IBA982101"="IITA-TMS-IBA982101",
                                             "TMS14F1159P003"="TMS14F1159P0003"))
Phenoleaf2 <- Phenoleaf1  %>% 
        group_by(accession_name) %>%
        dplyr::summarise_all(funs(mean(., na.rm = TRUE)))
Phenoleaf3 <- Phenoleaf2 %>%
               mutate(Clone = accession_name,
                      LeafArea = Total.Area) %>%
              select(Clone, LeafArea)

#complete with disease
PhenoDisea1 <- read_csv("data/IITA2021_Mich/PhenoDisea.csv")
spec(PhenoDisea1)
PhenoDisea2 <- PhenoDisea1 %>% 
           group_by(Clone) %>%
            dplyr::summarise_all(funs(mean(., na.rm = TRUE)))

#Phenocolen <- unique(Phenocol)

Phenosum1 <- left_join( PhenoDisea2, Phenocomp3)
Phenosum <- left_join(Phenosum1, Phenoleaf3)         
```

```{r}
AMMIgenoScores1 <- AMMIStatGXE$genoScores #using for genotype score
AMMIgenoScores2 <- as.data.frame(AMMIgenoScores1) 
AMMIgenoScores3 <- as_tibble(rownames_to_column(AMMIgenoScores2)) %>% 
                          mutate(Clone = rowname) %>%
                                 #genoScPC1 = PC1,
                                  #genoScPC2 = PC2) %>%
                         select(-rowname)

AMMIgenoMean1 <- AMMIStatGXE$genoMean #using for genotype mean (mean of root harvested by clone across all environment)
AMMIgenoMean2 <- as.data.frame(AMMIgenoMean1) 
AMMIgenoMean3 <- as_tibble(rownames_to_column(AMMIgenoMean2)) %>% 
                          mutate(Clone = rowname,
                                genoMean = AMMIgenoMean1) %>% 
                         select(-rowname, -AMMIgenoMean1)

StatGXEgeno1 <- left_join(AMMIgenoMean3, AMMIgenoScores3)
StatGXEgeno11 <- left_join(StatGXEgeno1, Phenosum)
StatGXEgeno11$branchP <- multMissing(StatGXEgeno11$branchP, maxIter = 10, naStrings = NULL)
StatGXEgeno11$LeafArea <- multMissing(StatGXEgeno11$LeafArea, maxIter = 10, naStrings = NULL)
StatGXEgeno2 <- as_tibble(StatGXEgeno11) %>% select(-Clone)
#jpeg(filename = "corrplogenocheck.jpeg", res = 400, width = 14, height = 8, units = "cm")
StatGXEgeno2_cor <- cor(StatGXEgeno2)
StatGXEgeno2_cor <- StatGXEgeno2_cor[1:4, ]
Figplot2 <- corrplot(StatGXEgeno2_cor, tl.cex = 0.5, number.cex=0.6, method = "number", sig.level = 0.01, insig = "blank")
#dev.off()  
```


##For the weather data by cycle time "old"
```{r}
StatGXEenv3 <- left_join(StatGXEenv1, weather_cycleT)
StatGXEenv4 <- as_tibble(StatGXEenv3) %>% select(-ENV, -ELEV, -LAT, -LONG, -RAIN, -SRAD, -TMIN, -TMAX, -RHUM, -PAR)
#corrplot(cor(StatGXEenv4), type = "upper")
StatGXEenvCOR <- StatGXEenv4 %>% summarise(TMIN1cor = cor(AMMIenvmean, TMIN1),
                    TMIN2cor = cor(AMMIenvmean, TMIN2),
                    TMIN3cor = cor(AMMIenvmean, TMIN3),
                    TMIN4cor = cor(AMMIenvmean, TMIN4),
                    TMAX1cor = cor(AMMIenvmean, TMAX1),
                    TMAX2cor = cor(AMMIenvmean, TMAX2),
                    TMAX3cor = cor(AMMIenvmean, TMAX3),
                    TMAX4cor = cor(AMMIenvmean, TMAX4),
                    SRAD1cor = cor(AMMIenvmean, SRAD1),
                    SRAD2cor = cor(AMMIenvmean, SRAD2),
                    SRAD3cor = cor(AMMIenvmean, SRAD3),
                    SRAD4cor = cor(AMMIenvmean, SRAD4),
                    RHUM1cor = cor(AMMIenvmean, RHUM1),
                    RHUM2cor = cor(AMMIenvmean, RHUM2),
                    RHUM3cor = cor(AMMIenvmean, RHUM3),
                    RHUM4cor = cor(AMMIenvmean, RHUM4),
                    RAIN1cor = cor(AMMIenvmean, RAIN1),
                    RAIN2cor = cor(AMMIenvmean, RAIN2),
                    RAIN3cor = cor(AMMIenvmean, RAIN3),
                    RAIN4cor = cor(AMMIenvmean, RAIN4))# %>% 
                
```



##For year 2017 specially with more than 30% missing values for 2017.
```{r}
#checks the number of clones (genotype) per environmeent then try to set the error message = "More than 30% missing values for 2017" Year 2017skipped.
EVAL2StatGXE12017 <- EVAL2StatGXE1 %>% dplyr::filter(year == 2017)
EVAL2StatGXE12017Ago <- EVAL2StatGXE12017 %>% filter(source == "Ago17")# 65 clones
EVAL2StatGXE12017Ibadan <- EVAL2StatGXE12017 %>% filter(source == "Ibadan17")#28
EVAL2StatGXE12017Ikenne <- EVAL2StatGXE12017 %>% filter(source == "Ikenne17")#34
EVAL2StatGXE12017Mokwa <- EVAL2StatGXE12017 %>% filter(source == "Mokwa17")#11
#filter EVAL2StatGXE12017Ag to use only the clones commons clones with the others locations
COL1 <- EVAL2StatGXE12017Ago$Clone
COL2 <- EVAL2StatGXE12017Ibadan$Clone
COL3 <- EVAL2StatGXE12017Ikenne$Clone
COL4 <- EVAL2StatGXE12017Mokwa$Clone
COL <- c(COL1 = COL3, COL1 = COL2, COL1 = COL4)
COLEN <- unique(COL)
COLENen <- data_frame(Clone = COLEN)
EVAL2StatGXE12017AgoFiltered <- left_join(COLENen, EVAL2StatGXE12017Ago)
EVAL2StatGXE22017 <- bind_rows(EVAL2StatGXE12017AgoFiltered, EVAL2StatGXE12017Ibadan, EVAL2StatGXE12017Ikenne, EVAL2StatGXE12017Mokwa)
EVAL2StatGXE2017 <- statgenSTA::createTD(data = EVAL2StatGXE22017, genotype = "Clone", trial = "source")
AMMIStatGXE2017 <- gxeAmmi(TD = EVAL2StatGXE2017, trait = "HWAMMS")
summary(AMMIStatGXE2017 )


AMMIenvscores1 <- AMMIStatGXE2017$envScores  
AMMIenvscores2 <- as.data.frame(AMMIenvscores1) 
AMMIenvscores3 <- as_tibble(rownames_to_column(AMMIenvscores2)) %>% 
                          mutate(env = rowname,
                                 envScoresPC1 = PC1,
                                  envScoresPC2 = PC2) %>%
                         select(-PC1, -PC2, -rowname)

EVAL2StatGXEenv <- EVAL2StatGXE1 %>% 
                     select(-Clone, -HWAMMS, -HWAMMSsq) %>%
                      mutate(env = source) %>%
                        dplyr::group_by(env) %>%
                         dplyr::summarise(ELEV = unique(ELEV), 
                                          LAT = unique(LAT), 
                                          LONG = unique(LONG), 
                                          RAIN = unique(RAIN), 
                                          SRAD = unique(SRAD), 
                                          TMIN = unique(TMIN), 
                                          TMAX = unique(TMAX),
                                          RHUM = unique(RHUM),
                                          PAR = unique(PAR))
StatGXEenv1 <- left_join(AMMIenvscores3, EVAL2StatGXEenv)
StatGXEenv2 <- as_tibble(StatGXEenv1) %>% select(-env)
corrplot(cor(StatGXEenv2[,-2]))
corrplot(cor(StatGXEenv2[,-1]))
corrplot(cor(StatGXEenv2))
```




## Perform a Finlay-Wilkinson analysis for all trials.
```{r}
MICHFW <- gxeFw(TD = EVAL2StatGXE, trait = "HWAMMS")
summary(MICHFW)
```


```{r}
#FWenvscores1 <- MICHFW$envEffs
FWenvscores1 <- MICHFW$envEffs %>%
            mutate(ENV = trial) %>%
                   select(ENV, envMean) 
#FWenvscores2 <- as_tibble(rownames_to_column(FWenvscores2)) %>% 
                       #   mutate(ENV = rowname,
                                 #envScoresPC1 = PC1,
                                  #envScoresPC2 = PC2) %>%
                              #  FWenvmean = FWenvscores1) %>% 
                         #select(-rowname, -FWenvscores1) #-PC1, -PC2, -rowname)
EVAL2StatGXEenv <- EVAL2StatGXE1 %>% 
                     select(-Clone, -HWAMMS, -HWAMMSsq) %>%
                      mutate(ENV = source) %>%
                        dplyr::group_by(ENV) %>%
                         dplyr::summarise(ELEV = unique(ELEV), 
                                          LAT = unique(LAT), 
                                          LONG = unique(LONG), 
                                          RAIN = unique(RAIN), 
                                          SRAD = unique(SRAD), 
                                          TMIN = unique(TMIN), 
                                          TMAX = unique(TMAX),
                                          RH2M = unique(RH2M),
                                          WIND = unique(WIND),
                                          PAR = unique(PAR))
StatGXEenv1 <- left_join(FWenvscores1, EVAL2StatGXEenv)
StatGXEenv2 <- as_tibble(StatGXEenv1) %>% select(-ENV)
#corrplot(cor(StatGXEenv2[,-2]))
#corrplot(cor(StatGXEenv2[,-1]))
corrplot(cor(StatGXEenv2))
```


```{r}
FWgenoScores1 <- MICHFW$estimates %>%
            mutate(Clone = genotype) %>%
                   select(Clone, genMean)
#AMMIgenoScores2 <- as.data.frame(AMMIgenoScores1) 
#AMMIgenoScores3 <- as_tibble(rownames_to_column(AMMIgenoScores2)) %>% 
                         # mutate(Clone = rowname,
                              #   genoScPC1 = PC1,
                              #    genoScPC2 = PC2) %>%
                         #select(-PC1, -PC2, -rowname)

StatGXEgeno1 <- left_join(FWgenoScores1, Phenosum)
StatGXEgeno1$branchP <- multMissing(StatGXEgeno1$branchP, maxIter = 10, naStrings = NULL)
StatGXEgeno1$LeafArea <- multMissing(StatGXEgeno1$LeafArea, maxIter = 10, naStrings = NULL)
StatGXEgeno2 <- as_tibble(StatGXEgeno1) %>% select(-Clone)
corrplot(cor(StatGXEgeno2))#, method = "number", type = "upper")
#corrplot(cor(StatGXEgeno2[,-2]))
#corrplot(cor(StatGXEgeno2[,-1]))

```


